\section{Quantum Network Master Equation}\label{C_quantum_network_master_equation}

The previous description for the noise in the quantum walk lacks a parameter that, in thermodynamics, correspond to the temperature. 
Thus, we propose another framework: instead of considering the jump operator as a transition between the nodes, we consider the thermal bath interaction in the energy states.
Let us retake the standard Lindblad equation \eqref{Lindbladian} with Hamiltonian $\hat H = \hat L$.
We introduce the basis $\{\ket{\lambda}\}_{\lambda}$ such that $\hat H = \sum_\lambda \lambda\ket{\lambda}\bra{\lambda} = \hat L$ is diagonal (the network must satisfy the detailed balance condition \eqref{detail_condition}).
We define the jump operator $\hat J_{\lambda\mu} = \ket{\lambda}\bra{\mu}$ as the jumps from the energy state $\ket{\mu}$ to the energy state $\ket{\lambda}$ obtaining the master equation
\begin{equation}\label{Lindblad_energy_jump}
    \frac{d}{dt}\hat\rho = -i\left[\hat L,\hat\rho\right] + \sum_{\lambda\mu} \gamma_{\lambda\mu} \left(\hat J_{\lambda\mu} \hat\rho \hat J^\dagger_{\lambda\mu} - \frac{1}{2}\left\{ \hat J^\dagger_{\lambda\mu}\hat J_{\lambda\mu}, \hat\rho\right\} \right),
\end{equation}
where the coefficients $\gamma_{\lambda\mu}$ indicate the probability of taking their respective jumps.

We assume that the dynamics will tend to a stationary distribution in the form of
\begin{equation}\label{pretended_stationary_distribution}
    \hat \rho^* = \frac{e^{-\beta\hat L}}{Z},
\end{equation}
with $Z = \Tr\left[e^{-\beta\hat L}\right]$ being the partition function.
The master equation for the stationary distribution \eqref{pretended_stationary_distribution} reduces to
\begin{equation}\label{cancel_master_equation}
    0 = -i\left[\hat H, \frac{e^{-\beta\hat L}}{Z}\right] + \sum_{\lambda\mu} \gamma_{\lambda\mu} \left(\hat J_{\lambda\mu}  \frac{e^{-\beta\hat L}}{Z} \hat J^\dagger_{\lambda\mu} - \frac{1}{2}\left\{ \hat J^\dagger_{\lambda\mu}\hat J_{\lambda\mu},  \frac{e^{-\beta\hat H}}{Z}\right\} \right).
\end{equation}
The first term on the r.h.s. vanishes since the commutator is zero.

Now, we analyze the dissipation terms independently.
The first one can be written as 
\begin{equation}\label{BCS_first_term}
        \sum_{\lambda\mu} \gamma_{\lambda\mu} \ket{\lambda}\bra{\mu}\frac{e^{-\beta\hat L}}{Z} \ket{\mu}\bra{\lambda} = 
        \sum_{\lambda\mu} \gamma_{\lambda\mu} \frac{e^{-\beta \epsilon_\mu}}{Z} \ket{\lambda}\bra{\lambda}.
\end{equation}
While the second becomes
\begin{equation}\label{BCS_second_term}
    \sum_{\lambda\mu} \gamma_{\lambda\mu}\left[\frac{1}{2}\ket{\mu}\braket{\lambda}{\lambda}\bra{\mu}\frac{e^{-\beta\hat L}}{Z} +\frac{1}{2}\frac{e^{-\beta\hat L}}{Z}\ket{\mu}\braket{\lambda}{\lambda}\bra{\mu}\right]= \sum_{\lambda\mu} \gamma_{\lambda\mu}\left[\frac{e^{-\beta \epsilon_\mu}}{Z}\ket{\mu}\bra{\mu}\right].
\end{equation}
Therefore, inserting the equations \eqref{BCS_first_term} and \eqref{BCS_second_term} into the master equation \eqref{cancel_master_equation}, we obtain the condition
\begin{equation}\label{Kirchhoff_law_energy}
    \sum_{\lambda\mu}\left[\gamma_{\lambda\mu}\frac{e^{-\beta \mu}}{Z} - \gamma_{\mu\lambda}\frac{e^{-\beta \lambda}}{Z}\right]\ket{\lambda}\bra{\lambda} = 0.
\end{equation}
This is the Kirchhoff's current law which says that the sum of all the currents must vanish. The system should satisfy this condition in order to have the Boltzmann distribution.
However, for a fixed $\beta$, there are several possible choices for the coefficients $\gamma_{\lambda\mu}$ such that equation \eqref{Kirchhoff_law_energy} holds.
Each different choice generates a different path to reach the stationary distribution \eqref{pretended_stationary_distribution}.
We are looking for a solution that is explicitly depends on the parameter $\beta$. The simplest choice is that the damping rates satisfies teh detailed balance condition 
%We assume that the system must satisfy the maximal entropy production principle: the system should relax to the stationary distribution following the trajectory that has maximal entropy production. The principle is satisfied when the Kirchhoff law \eqref{Kirchhoff_law_energy} reduces to the detail balance condition
\begin{equation}
    \gamma_{\lambda\mu}\frac{e^{-\beta \mu}}{Z} - \gamma_{\mu\lambda}\frac{e^{-\beta \lambda}}{Z} = 0,
\end{equation}
which has the solution
\begin{equation}\label{gamma_detailed_balance}
    \begin{split}
        \gamma_{\lambda\mu} = c \;e^{-\frac{\beta}{2}\left(\lambda - \mu\right)}\\
        \gamma_{\mu\lambda} = c \; e^{\frac{\beta}{2}\left(\lambda - \mu\right)}.
    \end{split}
\end{equation}
The solution is not unique;  there exist a set of possible solutions which differ by a constant $c$.
%with $Z_\gamma = \sum_{\lambda\gamma}e^{-\frac{\beta}{2}\left(\epsilon_\lambda - \epsilon_\mu\right)}$ is the renormalization factor.

The von Neumann entropy measure the mi

Taking the limit $\beta \rightarrow \infty$, that is $T  \rightarrow 0$, the transition rates tend to
\begin{equation}
    \gamma_{\lambda\mu} \rightarrow \left\{\begin{aligned}
        0 \qquad \lambda > \mu\\
        1 \qquad \lambda = \mu\\
        \infty \qquad  \lambda < \mu \\
    \end{aligned}\right. . 
\end{equation}
The transitions from lower to higher energy state are suppressed, while the opposite one are extremely favorite. Thus, the system is led to the zero energy state that is the stationary state
\begin{equation}\label{beta_inf_stationary_distribution}
    \hat\rho^* = \ket{\lambda = 0}\bra{\lambda = 0}.
\end{equation} 
It is a pure state therefore the von Neumann entropy vanishes.

In the opposite limit $\beta \rightarrow 0$, that is $T \rightarrow \infty$, the transition rates become
\begin{equation}
    \gamma_{\lambda\mu} \rightarrow 1.
\end{equation}
We have the opposite effect, the particle can jump across the different energy state with uniform probability. Thus, the stationary distribution is the maximal entropy state, i.e. the uniform distribution
\begin{equation}
    \hat\rho^* = \frac{1}{N}\begin{pmatrix}
        1&&0\\
        &\ddots&\\
        0&&1\\
    \end{pmatrix}.
\end{equation}
It is a maximal entropy state with $S = \ln N$.

\subsection{Return to node's basis}
We go back to the position basis $\{\ket{i}\}_{i<N}$, where $\ket{i}$ indicates the particle in the node $i$.  
The jump operators can be expressed as
\begin{align}
    \hat J_{\lambda\mu} = \sum_{ij} \braket{i}{\lambda}\braket{\mu}{j}\hat J_{ij}\\
    \hat J_{\lambda\mu}^\dagger = \sum_{ij} \braket{j}{\mu}\braket{\lambda}{i}\hat J_{ij}^\dagger
\end{align}
where $\hat J_{ij} = \ket{i}\bra{j}$.
Thus, substituting these expressions into equation into the equation \eqref{Lindblad_energy_jump} we obtain
\begin{equation}\label{quantum_network_position}
    \frac{d}{dt}\hat\rho = -i\left[\hat H,\hat\rho\right] +\sum_{ijkl} \gamma_{ij;kl} \left(\hat J_{ij}\hat\rho \hat J_{kl}^\dagger - \frac{1}{2}\left\{ \hat J_{kl}^\dagger\hat J_{ij}, \hat\rho\right\} \right),
\end{equation}
where the damping coefficient are define as
\begin{equation}\label{gamma_position}
    \gamma_{ij;kl}= \sum_{\lambda\mu}\gamma_{\lambda\mu}\braket{i}{\lambda}\braket{\lambda}{k}\braket{l}{\mu}\braket{\mu}{j}.
\end{equation}
Unlike the energy basis, the damping rates in the position basis are no longer diagonal, making the master equation \eqref{quantum_network_position} appear more complex as in the the equation\eqref{general_Lindblad_equation}. Using the detailed balance condition \eqref{gamma_detailed_balance} in \eqref{gamma_position}, we rewrite the damping rates as:.
\begin{equation}
    \gamma_{ij;kl}= \sum_{\lambda\mu}e^{-\frac{\beta}{2}\left(\epsilon_\lambda - \epsilon_\mu\right)}\braket{i}{\lambda}\braket{\lambda}{k}\braket{l}{\mu}\braket{\mu}{j}.
\end{equation}
\begin{comment}
    The master equation for the stationary distribution \eqref{pretended_stationary_distribution} in position space is 
    \begin{equation}
    0 = \sum_{ijkl} \gamma_{ij;kl} \left(\hat J_{ij}  \frac{e^{-\beta\hat H}}{Z} \hat J^\dagger_{kl} - \frac{1}{2}\left\{ \hat J^\dagger_{ij}\hat J_{kl},  \frac{e^{-\beta\hat H}}{Z}\right\} \right).
\end{equation}
The first term reduces to
\begin{equation}
    \sum_{ijkl}\gamma_{ij;kl} \ket{i}\bra{j} \frac{e^{-\beta\hat H}}{Z} \ket{l}\bra{k} = \sum_{ijkl}\gamma_{ij;kl} \frac{e^{-\beta H_{jl}}}{Z}\ket{i}\bra{k}.
\end{equation}
The anticommutator reduces to
\begin{equation}
    \begin{split}
        \sum_{ijkl}  \frac{\gamma_{ij;kl}}{2}\left(\ket{j}\braket{i}{k}\bra{l} \frac{e^{-\beta\hat H}}{Z}+ \frac{e^{-\beta\hat H}}{Z}\ket{j}\braket{i}{k}\bra{l}\right) \\ 
        = \sum_{ijlm} \left(\frac{\gamma_{ij;il}}{2}\frac{e^{-\beta H_{lm}}}{Z} + \frac{\gamma_{il;im}}{2}\frac{e^{-\beta H_{jl}}}{Z}\right)\ket{j}\bra{m}
    \end{split}
\end{equation}
We have applied some algebra and exchange dummy indices.
Thus, combining all together we obtain
\begin{equation}
    \begin{split}
        \sum_{ijkl}\left(\gamma_{ij;kl} \frac{e^{-\beta H_{jl}}}{Z} - \frac{\gamma_{ji;jl}}{2} \frac{e^{-\beta H_{lk}}}{Z} - \frac{\gamma_{jl;jk}}{2}\frac{e^{-\beta H_{il}}}{Z}\right)\ket{i}\bra{k} = 0\\
        \sum_{ijkl}\left(\gamma_{(ik)\leftarrow(jl)} \frac{e^{-\beta H_{jl}}}{Z} - \frac{\gamma_{(jj)\leftarrow(il)}}{2} \frac{e^{-\beta H_{lk}}}{Z} - \frac{\gamma_{(jj)\leftarrow(lk)}}{2}\frac{e^{-\beta H_{il}}}{Z}\right)\ket{i}\bra{k} = 0\\
        %\sum_{ijl}\left(\gamma_{ij;il} \frac{e^{-\beta H_{jl}}}{Z} - \frac{\gamma_{ji;ji}}{2} \frac{e^{-\beta H_{li}}}{Z} - \frac{\gamma_{jl;ji}}{2}\frac{e^{-\beta H_{il}}}{Z}\right)\ket{i}\bra{i} = 0
    \end{split}
\end{equation}
It can be interpreted as a Kirchhoff law in the link. Let suppose that the transition of the particle through the link is not instantaneous, then at each movement the particle chooses one link to travel. The probability to be in a specific link $(i,j)$ is given by $\frac{e^{-\beta H_{ij}}}{Z}$. In order to have a stationary distribution the dumping rate must satisfy the Kirchhoff law over the links distribution.
\end{comment}

In the high temperature limit, $\beta \rightarrow 0$, the damping rates are
\begin{equation}
    \gamma_{ij;kl}= \sum_{\lambda\mu} 1 \braket{i}{\lambda}\braket{\lambda}{k}\braket{l}{\mu}\braket{\mu}{j}.
\end{equation}
Using the completeness relation, we obtain
\begin{equation}
    \gamma_{ij;kl} = 1 \delta_{ik}\delta_{jl},
\end{equation}
where $\delta_{ik}$ is the Kronecker delta. Thus, the particle travel always through the same link.
In this case the position quantum network master equation \eqref{quantum_network_position} acquires a “symmetric" form
\begin{equation}
    \frac{d}{dt}\hat\rho = -i\left[\hat H,\hat\rho\right] +\sum_{ij}\left(\hat J_{ij}\hat\rho \hat J_{ij}^\dagger - \frac{1}{2}\left\{ \hat J_{ij}^\dagger\hat J_{ij}, \hat\rho\right\} \right).
\end{equation}
Since the relation \eqref{gamma_detailed_balance} can be modify by a constant factor, we recover the master equation for the Quantum Stochastic Walk \eqref{stochastic_lindblad_master}.
The stationary distribution is 
\begin{equation}
    \hat\rho^* = \frac{1}{N}\begin{pmatrix}
        1&&0\\
        &\ddots&\\
        0&&1\\
    \end{pmatrix}.
\end{equation}

\begin{comment}
In contrast, in the low temperature limit, $\beta \rightarrow \infty$, the damping coefficient are
\begin{equation}
    \begin{split}
        \gamma_{ij;kl}&= \sum_{\lambda\mu}e^{-\frac{\beta}{2}\left(\epsilon_\lambda - \epsilon_\mu\right)}\braket{i}{\lambda}\braket{\lambda}{k}\braket{l}{\mu}\braket{\mu}{j}\Theta(\lambda-\mu)\\
        & + \sum_{\lambda\mu} 1\braket{i}{\lambda}\braket{\lambda}{k}\braket{l}{\mu}\braket{\mu}{j}\delta_{\lambda\mu}\\
        & + \sum_{\lambda\mu} e^{-\frac{\beta}{2}\left(\epsilon_\lambda - \epsilon_\mu\right)}\braket{i}{\lambda}\braket{\lambda}{k}\braket{l}{\mu}\braket{\mu}{j}\Theta(\mu-\lambda)
    \end{split}
\end{equation}
The first term cancel out,  and the second is a sum of Kronecker delta. Thus, it reduces to
\begin{equation}\label{gamma_T=0}
    \begin{split}
        \gamma_{ij;kl}&= \sum_{\lambda}\braket{i}{\lambda}\braket{\lambda}{k}\braket{l}{\lambda}\braket{\lambda}{j}+ \sum_{\mu>\lambda}\infty\braket{i}{\lambda}\braket{\lambda}{k}\braket{l}{\mu}\braket{\mu}{j} \rightarrow \infty
    \end{split}
\end{equation}
\end{comment}
In contrast, in the low temperature limit, $\beta \rightarrow \infty$, the stationary distribution \eqref{beta_inf_stationary_distribution} for the node reduces to
\begin{equation}
    \hat\rho^* = \sum_{ij} \braket{i}{\lambda = 0}\braket{\lambda = 0}{j} \ket{i}\bra{j} = \sqrt{\rho^*_i\rho^*_j} \ket{i}\bra{j}
\end{equation}
The von Neumann entropy vanishes, indicating a pure state. 

%Collegare a overdamped case 


\begin{comment}
    \newpage
    The same distribution can be obtain in position space. As a matter of fact, the master equation \eqref{stochastic_lindblad_master} for the distribution \eqref{pretended_stationary_distribution}. Let assume that it is the stationary distribution, thus
    \begin{equation}
    0 = \sum_{\lambda\mu} \gamma_{ij} \left(\hat J_{ij}  \frac{e^{-\beta\hat H}}{Z} \hat J^\dagger_{ij} - \frac{1}{2}\left\{ \hat J^\dagger_{ij}\hat J_{ij},  \frac{e^{-\beta\hat H}}{Z}\right\} \right).
\end{equation}

the first term can be written as
\begin{equation}
\sum_{ij} \gamma_{ij} \ket{i}\bra{j}  \frac{e^{-\beta\hat H}}{Z} \ket{j}\bra{i} = 
\sum_{ji} \gamma_{ij} \frac{e^{-\beta H_{jj}}}{Z} \ket{i}\bra{i}
\end{equation}

Whereas, the second term becomes
\begin{equation}
\sum_{ij}\frac{\gamma_{ij}}{2}\ket{j}\bra{j} \frac{e^{-\beta\hat H}}{Z} + \frac{\gamma_{ij}}{2}\frac{e^{-\beta\hat H}}{Z}\ket{j}\bra{j}=
\sum_{ijk} \frac{1}{2}\left(\gamma_{ij}\frac{e^{-\beta H_{jk}}}{Z} + \gamma_{ik}\frac{e^{-\beta H_{jk}}}{Z}\right)\ket{j}\bra{k}
\end{equation}

We can combine together all the terms obtaining
\begin{equation}
\begin{split}
\sum_{ji} \left(\gamma_{ji}\frac{e^{-\beta H_{ii}}}{Z} - \gamma_{ij}\frac{e^{-\beta H_{jj}}}{Z}\right)\ket{j}\bra{j} = 0\\
\sum_{jk} \frac{1}{2}\left(\sum_i\gamma_{ij} + \sum_i\gamma_{ik}   \right)\frac{e^{-\beta H_{jk}}}{Z}\ket{j}\bra{k} = 0\\
\end{split}
\end{equation}
The first equation is again the Kirchhoff's current law, while the second is a condition over the parameter $\gamma$.

If the previous condition are satisfy the stationary distribution of the system is the canonical one. In our case it becomes
\begin{equation}
\hat\rho^* = \frac{e^{-\beta\hat L}}{Z} \qquad Z = \Tr[e^{-\beta\hat L}]  
\end{equation}
that is the density matrix \eqref{density_matrix} introduced by De Domenico to identify networks.
\end{comment}


\subsection{Analogy with the Network's Entropy}

The equation \eqref{Lindblad_energy_jump} describes the evolution of a quantum walk in the presence of classical noise. The temperature $T$ determines the noise strength.
Increasing the parameter $\beta$ suppresses the energy state with high eigenvalue, until just the zero eigenstate remains. Thus, the parameter $\beta$ allows us to analyze the information's spread along paths of chosen eigenstate. The von Neumann entropy measures the uncertainty over the state of the particle. 

This entropy behaves as the network's entropy \eqref{entropy}. 
The network's entropy measures the complexity of the spread of information across the network, which depends on the parameter $\beta$. 
In fact, for low values of $\beta$, all the possible channel are available, resulting in a complex spread of information and, thus, high entropy. As $\beta$ increases, the channels with high eigenvalue are suppressed until only the zero eigenstate remains, thus, zero entropy. 

In addition, we can apply a Wick rotation to the quantum dynamics \eqref{Lindblad_energy_jump} returning to a special random walk of classical particle. This rotation connect the stationary distribution at inverse temperature $\beta$ \eqref{pretended_stationary_distribution} with the propagator at time $t$ \eqref{random_walk_solution}.
\begin{equation}
e^{-\beta\hat L} \rightarrow e^{-t\hat L}
\end{equation}
Thus, cooling down the quantum system is analogous to the temporal evolution of the classical one. In fact, the two limits $\beta \rightarrow \infty$ and $t \rightarrow \infty$ converge to the same distribution: the system will be entirely in the zero eigenstate of the Laplacian.
Moreover, the density distribution \eqref{pretended_stationary_distribution} is always in the maximal entropy state. As a consequence, also the distribution for the classical random walk should cross state with maximal entropy.

The complexity of the possible paths is encoded in the von Neumann entropy as explained in chapter \ref{C_Density_Matrix}.
The entropy allows us to classify different networks based on the dynamical properties of the network itself. 
We can achieve it introducing the Kullback-Leibler divergence \eqref{KL_divergence} and the Jensen-Shannon divergence \eqref{JS_metric}.
However, because these quantities employ the trace of a Laplacian's function, the entropy studies only the spectral properties of the network. Therefore, networks with same spectrum but different structures and eigenstates may be indistinguishable using these methods.