\chapter{Quantum Network Master Equation}

In chapter \ref{C_Density_Matrix} we have introduce the concept of density matrix for a network, based on the communicability matrix. This quantity considers the correlation between the nodes created by the random walk dynamics. 

In this chapter we try to unify the two concept taken from the quantum realm: the quantum random walk and the density matrix.
In fact, we considering a quantum walk process exposed to thermal noise we obtain a stationaty distribution which coincides with the network's density matrix. 
The interaction between the quantum system and the thermal noise are considered Markovian, thus, it is studied with the Lindblad master equation.

\input{Quantum_Stochastic_Walk}

\input{BCS_quantum_network_master_equation}


\section{Generalization to other dynamic}

Until now, we have examined only the random walk on network, but this framework can be generalize to other more complex dynamics on network \cite{De_Domenico_2023}.
The dynamics should be linear such that the evolution of the observable per node $i$ are 
\begin{equation}\label{general_dynamics}
    \frac{d}{dt} x_i = \sum_j H_{ij} x_j,
\end{equation}
where $H_{ij}$ controls the evolution of the system.
For the continuos time random walk the control matrix coincides with the Laplacian.
In order to apply the Wick rotation and we obtain the quantum version of the system.
Let $\{\ket{i}\}_{i<N}$, $N$ is the number of node, be a orthonormal basis for the Hilbert space $\mathcal{H}$, the state $\ket{\psi}$ is defined as
\begin{equation}
    \ket{\psi}= \sum_i\sqrt{x_i}\ket{i}
\end{equation}
such that $x_i = |\braket{i}{\psi}|^2$.
The evolution follow the Schrödinger equation
\begin{equation}
    \frac{d}{dt}\ket{\psi(t)} = -i\hat H\ket{\psi(t)}
\end{equation}
where 
$\hat H = \sum_{ij} H_{ij} \ket{i}\bra{j}$ is the control operator or the chosen dynamics. To satisfy the Schrödinger equation the control operator must be symmetric.
Now we can add thermal noise arriving at quantum master equation \eqref{Lindblad_energy_jump} which has stationary distribution 
\begin{equation}
    \rho^*= \frac{1}{Z}e^{-\beta\hat H}
\end{equation}
with $Z= \Tr\left[e^{-\beta\hat H}\right]$ is the partition function.

Therefore, the network entropy for network under the dynamics \eqref{general_dynamics} is 
\begin{equation}
    S = -\Tr\left[\frac{1}{Z}e^{-\beta\hat H}\ln\left(\frac{1}{Z}e^{-\beta\hat H}\right) \right].
\end{equation}
Based on the considered dynamics the network will have a different entropy's value.



\section{Symmetry breaking}

Until now, we have considered the network holding the detail balance condition and, therefore, be mapped in a symmetric matrix; 
but the majority of the networks do not satisfy this condition. 

To deal with them, we modify slightly the Lindblad master equation \eqref{stochastic_lindblad_master}. 
As a matter of fact, in the chapter \ref{C_Lindblad} we have analyzed also the case where the interaction with the environment is not symmetric \eqref{environment_coefficients}. Thus, taking the dissipative part of the equation \eqref{C_rotating_wave} in the Schrödinger picture with the coefficients $\Gamma_{ij} = L_{ij}$ and the jump operators $J_{ij} = \ket{i}\bra{j}$ we obtain 
\begin{equation}
    \frac{d\hat\rho(t)}{dt} = \sum_{ij}\Gamma_{ij}\left[\hat J_j\hat\rho(t),\hat J_i^\dagger\right]+\Gamma_{ji}^\dagger\left[\hat J_j,\hat\rho(t)\hat J_i^\dagger\right].
\end{equation}
Isolating the symmetric and antisymmetric part of the Laplacian, respectively $\gamma_{ij} = \left(L_{ij} + L_{ji}\right)$ and $\pi_{ij} =  \frac{-i}{2}\left(L_{ij}-L_{ji}\right)$ such that $\Gamma_{ij}(\omega) =\frac{1}{2}\gamma_{ij}(\omega)+i\pi_{ij}(\omega)$, we arrive to the equation
\begin{equation}
    \frac{d\hat\rho(t)}{dt} = \sum_{ij}\gamma_{ij}\hat J_j\hat\rho(t)\hat J_i^\dagger -\frac{\gamma_{ij}}{2}\left\{\hat J_i^\dagger\hat J_j,\hat\rho(t)\right\} + i\pi_{ij}\left[\hat J_i^\dagger\hat J_j,\hat\rho(t)\right],
\end{equation}
where $[\cdot,\cdot]$ and $\{\cdot,\cdot\}$ are respectively the commutator and anticommutator.

Let define a new Hamiltonian $\hat H_{A} = \sum_{ij}\pi_{ij}\hat J_i^\dagger\hat J_j$ that encodes the dynamics of the not symmetric part. 
It give origin to a coherent dynamics that follow the Von Neumann equation. As a matter of fact the total dynamics can be written as
\begin{equation}\label{antisymmetric_master_equation}
    \frac{d\hat\rho(t)}{dt} = i\left[\hat H_{A},\hat\rho(t)\right] + \sum_{ij}\gamma_{ij}\hat J_j\hat\rho(t)\hat J_i^\dagger -\frac{\gamma_{ij}}{2}\left\{\hat J_i^\dagger\hat J_j,\hat\rho(t)\right\}.
\end{equation}

The dynamics \eqref{antisymmetric_master_equation} does not converge no more to a stationary state due to the Von Neumann part.
We can generalize as in the \eqref{stochastic_lindblad_master} 

\begin{equation}
    \frac{d}{dt}\hat \rho = -i\left[\hat H + \omega\hat H_{A},\hat\rho\right] + \sum_{ij}\gamma_{ij}\left[\hat J_{ij} \hat\rho\hat J_{ij}^\dagger -\frac{1}{2} \left\{ \hat J_{ij}^\dagger \hat J_{ij}, \hat\rho\right\}\right].
\end{equation}
where $\hat H$ is the hermitian part of the Laplacian operator.


\bigskip

The quantum walk on network with a not hermitian Hamiltonian has a evolution operator that is not unitary. It can be split between the hermitian $\hat H_S$ and anti-hermitian $\hat H_A$ components as
\begin{equation}\label{not_hermitian_time_operator}
    U(t,0) = e^{-it\hat H_S} e^{-t\hat H_A}
\end{equation}
with
\begin{equation}
    \hat H_S =  \frac{1}{2}\left(\hat H + \hat H^\dagger\right) \qquad \qquad
    \hat H_A = \frac{-i}{2}\left(\hat H - \hat H^\dagger\right)
\end{equation}

If the Hamiltonian $\hat H$ is positive defined, the second exponential of the equation \eqref{not_hermitian_time_operator} dissipate energy. 
Thus, the system can be studied again with a Lindblad master equation \eqref{Lindbladian} in the form
\begin{equation}
    \frac{d}{dt}\hat\rho = -i\left[\hat H_S,\hat\rho\right] + \sum_k \gamma_k \left(\hat J_k \hat\rho \hat J^\dagger_k - \frac{1}{2}\left\{ \hat J^\dagger_k\hat J_k, \hat\rho\right\} \right).
\end{equation}
The jumps operators must reproduce the same dissipation as the anti-hermitian operator $\hat H_A$.

The damping coefficients do not have to hold the Kirchhoff law. Thus, the system may not converge to a stationary state but may persist some stationary density currents.