\chapter{Introduction to Network Theory}\label{Network_Theory}

%Networks are all around us, shaping the very fabric of our world. 
From social interactions and transportation systems to the intricate connections within biological organisms and the internet, networks provide a powerful framework to understand complex systems. Graph theory, the mathematical foundation of network theory, offers the tools needed to analyze these interconnected structures.

In this chapter, we introduce the fundamental concepts of network and graph theory, along with various types of random graphs.
Then, we focus on random walks and diffusion processes on networks, considering both classical and quantum cases, which play a crucial role in modeling real-world phenomena such as information spread, epidemic modeling, and quantum transport.


\section{Introduction to Graph Theory}

%definition of network, adjacency matrix, laplacian matrix, shortest path, ...
The mathematical framework used in Network theory is given by Graph theory. 

A graph is defined by an ordinate couple $(V,E)$ where $V = \{1,2,3, ...,n\}$ is the set of nodes or vertices and $E = \left\{ (i, j): i , j \in V ; i \text{ is linked to } j\right\}$ is the set of links or edges. Usually, a general graph is denoted as $G =(N,M)$ where $N$ and $M$ are the cardinality of $V$ and $E$ respectively.

A graph can be viewed also via a matrix called Adjacency matrix which is defined as
\begin{equation}
    A_{ij}= \left\{ \begin{aligned}
        +1 &\quad \mathrm{if} \; i \; \mathrm{is ~linked ~to} \; j \\
        0 &\quad \mathrm{otherwise} \\
    \end{aligned} \right.  .
\end{equation}
The degree $d_i$ of a node $i$ is the number of nodes to which it is connected. We can introduce the degree matrix $D$ as $D_{ij} = d_i\delta_{ij}$
It can be computed from the adjacency matrix as
\begin{equation}
    d_i = \sum_j A_{ij}.
\end{equation}

Graphs can be grouped mainly into two type: \textit{undirected} and \textit{directed} graph. In the first one if the node $i$ is linked to $j$ then $j$ is linked to $i$, namely $(i, j) = (i, j)$; its adjacency matrix is symmetric.
In contrast, in the second one, if the node $i$ is linked to $j$ not necessary $j$ is linked to $i$, namely $(i, j) \neq (i, j)$; its adjacency matrix is not symmetric.

An important concept in graph theory is the study of connections between nodes that are not directly linked by an edge. As a matter of fact, two nodes can be connected by passing through multiple other nodes.
A \textit{walk} of length $k$ from node $i$ to node $j$ is a sequence of nodes $(x_0,x_1,...,x_k)$ such that $x_0=i$, $x_k=j$ and $(x_l, x_{l+1}) \in E$ for all $l \in \{0, ..., k-1\}$. A node can be crossed multiple times.  
If a walk visits each node only once, it is called a \textit{path}.
A particularly important concept is the \textit{shortest path} or \textit{geodesic} that is the path that crosses the minimum number of nodes.  
The number of walks $W_{ij}(k)$ of length $k$ from node $i$ to node $j$ can be computed using the adjacency matrix as
\begin{equation}
    W_{ij}(k) = (A^k)_{ij} .
\end{equation}

A graph is said to be \textit{connected} if, for each pair of distinct nodes $i$ and $j$, there exists a walk that connects them. 
We can defined $G' =(V',E')$ a subgraph of $G= (V,E)$ if $V' \subseteq V$ and $E' \subseteq E$.
A \textit{component} of a graph $G= (V,E)$ is a connected subgraph $G' =(V',E')$ meaning that not connected to any external node of the graph, that is $(i,j)\notin E$ for each $i\in V'$ and $j\in V\setminus V'$. 
A important concept is the \textit{giant component}: a connected subgraph that has approximately the same number of nodes of the total graph. 

\input{random_Network.tex}

\input{Random_walk.tex}

\input{Quantum_Walk.tex}