\subsection{Network of quantum harmonic oscillators}

If we consider a system as quantum in place of a classic one the covariance matrix changes and becomes similar to te exponential communicability matrix.
In the quantum case, the quantities $H_L$, $q_i$ and $p_j$ are promoted to operators $\hat H_L$, $\hat q_i$ and $\hat p_j$ and they satisfy the commutator relation $\left[\hat q_i, \hat p_j\right] = i \delta_{ij}$ ($\hbar = 1$).

In order to study the system we need to add a new term; it should be considered as additional springs with constant $K'$ that connect each node to the ground: it prevent the the center of mass from moving. Thus, the Hamiltonian becomes
\begin{equation}\label{H_L_QM}
    \hat H_L = \sum_i\left(\frac{\hat p_i^2}{2}+\frac{K'}{2}\hat q_i^2\right) + \sum_{ij}\frac{1}{2} L_{ij}\hat q_i\hat q_j.
\end{equation}

We introduce the creation and annihilation operators as
\begin{equation}
     \hat a_i = \frac{1}{\sqrt{2}}\left(\sqrt{\Omega} \hat q_i + \frac{i}{\sqrt{\Omega}}\hat p_i\right) \qquad 
     \hat a_i^\dagger = \frac{1}{\sqrt{2}}\left(\sqrt{\Omega} \hat q_i - \frac{i}{\sqrt{\Omega}}\hat p_i\right), 
\end{equation}
and the inverse as
\begin{equation}
    \hat q_i = \sqrt{\frac{1}{2\Omega}}\left(\hat a_i + \hat a_i^\dagger\right) \qquad
    \hat p_i = i\sqrt{\frac{\Omega}{2}}\left(\hat a_i - \hat a_i^\dagger\right),
\end{equation}
where $\Omega = \sqrt{K'}$.
They satisfy the commutation relation $\left[\hat a_i, \hat a^\dagger_j\right] = \delta_{ij}$. 

The Hamiltonian can be written as 
\begin{equation}
    \hat H_L = \sum_i \Omega \left(\hat a_i\hat a^\dagger_i + \frac{1}{2}\right) + \frac{1}{4\Omega}\sum_{ij}\left(\hat a_i +\hat a_i^\dagger\right)L_{ij}\left(\hat a_i +\hat a_i^\dagger\right).
\end{equation}

Assuming the network satisfies the detailed balance condition \eqref{detail_condition}, $L$ is symmetric and, therefore, we can diagonalize it. The diagonalized laplacian is written in the form $\Lambda = OLO^T$, where $O$ is a orthogonal matrix.
This generates a new pair of creation and annihilation operators respect the eigenvalue $\mu$ of the Laplacian
\begin{equation}
    b_\mu = \sum_j a_jO_{\mu j}  \qquad \hat b_\mu^\dagger = \sum_j a_j^\dagger O^T_{\mu j} .
\end{equation}

Thus, the new Hamiltonian becomes a sum of independent Hamiltonians
\begin{equation}
    \hat H_L = \sum_\mu \hat H_\mu,
\end{equation}
with
\begin{equation}
    \begin{split}
        \hat H_\mu &= \Omega \left(\hat b_\mu\hat b^\dagger_\mu + \frac{1}{2}\right) + \frac{1}{4\Omega}\mu\left(\hat b_\mu +\hat b_\mu^\dagger\right)^2.\\
        %&= \Omega \left(\hat b_\mu\hat b^\dagger_\mu + \frac{1}{2}\right) + \frac{1}{4\Omega}\mu\left[\left(\hat b_\mu\right)^2 +\left(\hat b_\mu^\dagger\right)^2 + 2 \hat b_\mu \hat b_\mu^\dagger + 1 \right]\\
        %&= \Omega \left[ 1 + \frac{1}{2\Omega}\mu\right] \left(\hat b_\mu\hat b^\dagger_\mu + \frac{1}{2}\right) + \frac{1}{4\Omega}\mu\left[\left(\hat b_\mu\right)^2 +\left(\hat b_\mu^\dagger\right)^2 \right].
    \end{split}
\end{equation}
With some algebra, it reduces to 
\begin{equation}
    \hat H_\mu = \Omega \left[ 1 + \frac{1}{2\Omega}\mu\right] \left(\hat b_\mu\hat b^\dagger_\mu + \frac{1}{2}\right) + \frac{1}{4\Omega}\mu\left[\left(\hat b_\mu\right)^2 +\left(\hat b_\mu^\dagger\right)^2 \right].
\end{equation}


We now consider the system as fermionic so the modes do not excite beyond the first excitation state. As a consequence, we can restrict the Hilbert state of a mode to the span of the ground state $\ket{g}$ and the first excited state $\ket{e_\mu}=b_\mu^\dagger\ket{g}$. Therefore, the second term in the Hamiltonian cancel out. 

Now, we can compute the thermal Green function or Matsubara Green function for fermions. This quantity describes the probability amplitude for the particle to travel from one state to the other in a given time $\tau$ (more details in the Appendix \ref{Appendix_B}).
In the Heisenberg picture, the annihilation and creation operator depend on time \cite{Coleman_2015} as
\begin{equation}
    \hat b_\mu (\tau) = e^{-\mu t} \hat b_\mu \qquad \hat b_\mu^\dagger (\tau) = e^{\mu t} 
\end{equation}
For $\tau > 0$ it is
\begin{equation}\label{Thermal_GReen_Function}
    \begin{split}
        G^L_{ij}(\beta, \tau > 0) &= \frac{\Tr\left[e^{-\beta \hat H_L}\hat a_i (\tau)  \hat a_j^\dagger\right]}{\Tr\left[e^{-\beta \hat H_L}\right]} =\sum_{\mu\nu} O_{\mu i}\frac{\Tr\left[ e^{-\beta \hat H_L} \hat b_\mu (\tau)\hat b_\nu^\dagger\right]}{\Tr\left[e^{-\beta \hat H_L}\right]}O_{j\nu}\\
        %&= \sum_{\mu} O_{i\mu}\left\{-e^{-\mu \tau}\left[\left(1-f\left( \Omega + \frac{1}{2\Omega^2}\mu\right)\right)\Theta(\tau)\right]\right\}O_{j\mu}\\
        %&= \sum_{\mu} O_{i\mu}\left\{\frac{e^{-\mu \tau}}{e^{-\beta\left[ \Omega + \frac{1}{2\Omega^2}\mu\right]} + 1}\right\}O_{j\mu}\\
    \end{split}
\end{equation}
The equation \eqref{Thermal_GReen_Function} reduces, the calculation are in the Appendix \ref{Appendix_B}, to
\begin{equation}
    G^L_{ij}(\beta, \tau > 0) = \sum_{\mu} O_{i\mu}\left\{-e^{-\mu \tau}\left[\left(1-f\left( \Omega + \frac{1}{2\Omega^2}\mu\right)\right)\right]\right\}O_{j\mu},
\end{equation} 
where $f(\epsilon)$ is the Fermi-Dirac distribution
\begin{equation}
    f(\epsilon)=\frac{1}{e^{\beta\epsilon} + 1}.
\end{equation}
With some algebra we obtain
\begin{equation}
    G^L_{ij}(\beta, \tau > 0) = \sum_{\mu} O_{i\mu}\left\{\frac{e^{-\mu \tau}}{e^{-\beta\left[ \Omega + \frac{1}{2\Omega^2}\mu\right]} + 1}\right\}O_{j\mu}
\end{equation}
In the limit $\tau \rightarrow 0^+$ and $\beta$ large enough it tend to
\begin{equation}\label{Quantum_correlation}
    G^L(\beta) = \sum_{\mu} O_{i\mu}{e^{\beta\left[ \Omega + \frac{1}{2\Omega^2}\mu\right]}} O_{\mu i},
\end{equation}
that can be written also as
\begin{equation}\label{quantum_correlation_2}
    G^L_{ij}(\beta)= e^{\beta \Omega}e^{\frac{\beta}{2\Omega}L}.
\end{equation}

Redefining the temperature $\beta \rightarrow \frac{\beta}{2\Omega}$ the equations \eqref{quantum_correlation_2} is related to equation \eqref{Estrada indeces}
\begin{equation}
    G^E(L) = e^{-\beta\Omega}G^L(\beta).
\end{equation}
When the temperature $T \rightarrow 0$ the communicability between the nodes drops to zero and the perturbation does not spread across the network. Instead, when $T \rightarrow \infty$ the communicability tend to infinity and the perturbation spread instantaneously.  

In the high temperature limit, $\beta \rightarrow 0$, the correlation of the quantum system \eqref{Quantum_correlation} converges to the correlation of the classic one \eqref{classic_correlation}. This is well known, since in the high temperature limit the quantum effects should be negligible.
